SAM simétrico y asimétrico

### 2020-06-25 analysis/paper/paper.Rmd
Adds some figures

---
title: "Title Goes Here"
author:
  - Myfirstname Mylastname:
      email: fl@oneplace.edu
      institute: [UofO]
      correspondence: true
  - Yourfirstname Yourlastname:
      email: fl@another.edu
      institute: [UofA]
      correspondence: false
institute:
  - UofO: University of One Place
  - UofA: University of Another Place
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

@@

```{r setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300
)
```

@@

```{r}
library(asymsam)
library(ggplot2)
```

```{r}
ggplot() +
  geom_qmap()
```

Here is a citation [@Marwick2017]

# Background

@@

@@

<!-- Here's some example analysis code: -->

```{r get-data, eval = FALSE}
# Note the path that we need to use to access our data files when rendering this document
my_data <- read.csv(here::here('analysis/data/raw_data/my_csv_file.csv'))
```

```{r demo-plot, fig.cap="A plot of random numbers"}
plot(rnorm(10))
```

Figure \@ref(fig:demo-plot) shows how we can have a caption and cross-reference for a plot

```{r demo-inline-code}
x <- round(pi, 2)
```

>>>>

---
title: "SAM simétrico y asimétrico"
author: "Elio Campitelli"
output: 
   # pdf_document:
   #    latex_engine: xelatex
   powerpoint_presentation:
    # ioslides_presentation:
        fig_height: 7.6
        fig_width: 12.8
        reference_doc: "../templates/template.pptx" 
        slide_level: 2
urlcolor: blue
header_includes:
   - \usepackage[utf8]{inputenc}
---

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300
)

library(metR)
library(data.table)
library(ggplot2)
library(magrittr)
library(ggh4x)
library(ggperiodic)
library(asymsam)

# source(here("scripts", "helperfun.R"))
# source(here("scripts", "data_locations.R"))
theme_set(theme_asymsam())
ZeroBreaks <- AnchorBreaks(0, NULL, 0)

lev.lab <- AddSuffix(" hPa")

guide_colorstrip_bottom <- function(width = 25, height = 0.5, ...) {
  guide_colorstrip(title.position = "top", title.hjust = 0.5,
                   barheight = height,
                   barwidth = width, ...)
}

main_levs <- c(700, 300, 30)
# knitr_set_timer()
# knitr_set_cache(cache.extra = 44)
```

```{r read-hgt, cache = FALSE}
hgt <- ReadNetCDF(ERA5(), 
                  subset = list(time = c("1979-01-01", "2018-12-31"),
                                latitude = c(-90, 10)),
                  vars = c(hgt = "z")) %>% 
  normalise_coords() %>% 
  .[, hgt := hgt/9.8] %>% 
  .[, dataset := "ERA5"] %>%  
  .[, hgt_a := hgt - mean(hgt), by = .(lon, lat, lev, month(time), dataset)] %>% 
  .[, hgt_m := mean(hgt_a), by = .(lat, lev, time)] %>% 
  .[, hgt_z := hgt_a - hgt_m]
```

```{r compute-SAM}
SAM <- hgt[lat <= -20] %>% 
  .[, hgt := hgt_a*sqrt(cos(lat*pi/180))] %>% 
  .[, .(eof = list(eof_flip(EOF(hgt ~ time | lon + lat, n = 1, data = .SD)))), 
    by = .(dataset, lev)]
```

@@

@@

```{r method, fig.cap = "Spatial patterns of the first EOF of 700 hPa geopotential height. Full field (left), zonally asymmetric component (middle) and zonally symmetric component (right). Arbitrary units."}
sam_sep <- SAM[, eof[[1]]$right, by = .(dataset, lev)] %>% 
  setnames("hgt", "full") %>% 
  .[, c("sym", "asym") := list(mean(full), Anomaly(full)), by = .(lat, lev, dataset)]

lab_sam <-  c(full = "Full", 
              asym = "Asymmetric",
              sym  = "Symmetric")

sam_sep %>% 
  .[dataset == "ERA5"] %>% 
  .[lev == 700] %>%
  rm_singleton() %>% 
  melt(id.vars = c("lon", "lat")) %>%

  # .[variable == "asym", value := value + rnorm(.N, 0, sd = 1e-6)] %>%
  .[, variable := factor(variable, levels = names(lab_sam), ordered = TRUE)] %>% 
  periodic(lon = c(0, 360)) %>% 
  ggplot(aes(lon, lat)) +
  # geom_contour2(aes(z = value, linetype = factor(-sign(..level..))), global.breaks = FALSE) +
  geom_contour_fill(aes(z = value), global.breaks = FALSE) +
  geom_contour_tanaka(aes(z = value), global.breaks = FALSE) +
  geom_qmap() +
  scale_x_longitude() +
  scale_y_latitude(limits = c(NA, -20)) +
  scale_fill_divergent(guide = "none") +
  coord_polar() +
  facet_grid(~variable, labeller = labeller(variable = lab_sam, lev = lev.lab)) 
```

@@

```{r compute-asymsam}
indexes <-  hgt[lat <= -20, eof_asym(hgt_a, lon, lat, time, n = 1), by = lev]
indexes_wide <- dcast(indexes, + lev + time ~ term, value.var = "estimate_norm")
```

```{r asymsam-timeseries, fig.cap = "Time series for the asymmetric SAM and symmetric SAM."}
indexes %>% 
  .[lev == 700 & term != "full"] %>% 
  ggplot(aes(time, estimate_norm)) +
  geom_line(aes(color = term)) +
  # geom_smooth(method = "lm", aes(color = term)) +
  scale_color_brewer(NULL, palette = "Set1", labels = lab_sam) +
  scale_y_continuous(NULL) +
  scale_x_datetime(NULL)
```

```{r}
indexes_wide[, cor(asym, sym), by = lev] %>% 
  ggplot(aes(lev, V1)) +
  geom_line() +
  scale_x_level() + 
  scale_y_continuous(limits = c(0, 1)) +
  coord_flip()
```

```{r}

indexes_2d <- hgt[lev %in% main_levs] %>% 
  indexes_wide[., on = .NATURAL, allow.cartesian = TRUE] %>% 
  .[, rbind(as.data.table(FitLm(hgt_a, sym, asym)),
            as.data.table(FitLm(hgt_a, full))),
    by = .(lon, lat, lev)] %>% 
  rm_intercept() %>% 
  periodic(lon = c(0, 360)) %>% 
  .[, term := factor(term, levels = names(lab_sam), ordered = TRUE)]

indexes_2d %>% 
  .[lat <= -20] %>% 
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = estimate), breaks = ZeroBreaks, global.breaks = FALSE) +
  # geom_contour2(aes(z = estimate, linetype = factor(-sign(..level..))), 
                # size = 0.3,
                # global.breaks = FALSE) +
  geom_contour_tanaka(aes(z = estimate),
                      breaks = ZeroBreaks,
                      size = 0.3,
                      global.breaks = FALSE) +
  geom_qmap(~.x[lat <= -20]) +
  # geom_contour2(data = ~.x[dataset == "NCEP"], aes(z = V1)) +
  scale_x_longitude() +
  scale_y_latitude(limits = c(NA, -20)) +
  scale_fill_divergent(NULL, guide = guide_colorstrip_bottom(), 
                       limits = c(-80, 80),
                       oob = scales::squish,
                       breaks = MakeBreaks(binwidth = 20, exclude = 0)) +
  scale_linetype(guide = "none") +
  coord_polar() +
  facet_grid(term ~ lev, labeller = labeller(term = lab_sam, lev = lev.lab))
```
